### 2016-03-04 paper/icfp16-submission/interactive.tex
hide interactive semantics and exploration figures

\subsection{Tracing Semantics}
\label{sec:inter-semant}
%
\begin{figure*}[t]
%% \relDescription{Computing Sub-Terms}
%% \begin{gather*}
%% \begin{array}{lcl}
%% \subtermssym                 & \dcolon & e \to \tr \\
%% \subterms{\eapp{e_1}{e_2}}   & \defeq & \subterm{\eapp{e_1}{e_2}}{e_1}; \subterm{\eapp{e_1}{e_2}}{e_2} \\
%% \subterms{\eplus{e_1}{e_2}}   & \defeq & \subterm{\eplus{e_1}{e_2}}{e_1}; \subterm{\eplus{e_1}{e_2}}{e_2} \\
%% \subterms{\eif{e_1}{e_2}{e_3}}   & \defeq & \subterm{\eif{e_1}{e_2}{e_3}}{e_1}; \\
                                %% &        & \subterm{\eif{e_1}{e_2}{e_3}}{e_2}; \\
                                %% &        & \subterm{\eif{e_1}{e_2}{e_3}}{e_3} \\
%% \subterms{\elet{x}{e_1}{e_2}}   & \defeq & \subterm{\elet{x}{e_1}{e_2}}{e_1}; \\
                                %% &        & \subterm{\elet{x}{e_1}{e_2}}{e_2} \\
%% \subterms{\efun{x}{e}}       & \defeq & \subterm{\efun{x}{e}}{e} \\
%% \subterms{e}                 & \defeq & \bullet
%% \end{array}
%% \end{gather*}
\judgementHead{Traced Evaluation}{\stepg{e}{\su}{\tr}{e}{\su}{\tr}}
\begin{gather*}
\inference[\recontext]
  {\stepg{e}{\su}{\tr}{e'}{\su'}{\tr'}}
  {\stepg{C[e]}{\su}{\tr}{C[e']}{\su'}{\singlestep{C[e]}{C[e']}; \tr'}}
\\ \\
\inference[\reappgood]
  {\pair{\efun{x}{e}}{\su'} = \force{v_1}{\tfun{\thole{}}{\thole{}}}{\su}}
  {\stepg{\eapp{v_1}{v_2}}{\su}{\tr}
         {e\sub{x}{v_2}}{\su'}{\singlestep{\eapp{v_1}{v_2}}{e\sub{x}{v_2}}; \tr}}
\\ \\
\inference[\reappbad]
  {\pair{\stuck}{\su'} = \force{v_1}{\tfun{\thole{}}{\thole{}}}{\su}}
  {\stepg{\eapp{v_1}{v_2}}{\su}{\tr}{\stuck}{\su'}{\singlestep{\eapp{v_1}{v_2}}{\stuck}; \tr}}
\end{gather*}
\caption{A selection of the operational semantics from
  Figure~\ref{fig:operational}, extended to collect a full reduction
  graph.}
\label{fig:interactive}
\end{figure*}

@@

\paragraph{Tracing Semantics}
%
We extend the transition relation (\S~\ref{sec:semantics}) to
collect the set of edges corresponding to the reduction graph.
%
Concretely, we extend the operational semantics to
a relation of the form $\stepg{e}{\vsu}{\tr}{e'}{\vsu'}{\tr'}$
where $\tr'$ collects the edges of the transition.

\paragraph{Collecting Edges}
%
Next, we describe the general recipe for extending a transition
rule to collect edges.
%
The general recipe for collecting steps-to edges is to
to recording the consequent of each original rule in the
trace. That is, each original judgment \step{e}{\vsu}{e'}{\vsu'}
becomes \stepg{e}{\vsu}{\tr}{e'}{\vsu'}{\singlestep{e}{e'}; \tr}.
%
As the translation is mechanical, we provide a selection of examples
in Figure~\ref{fig:interactive}.

@@

Next, we show how to build a visual interactive debugger
from the traced semantics, by describing the visualization
\emph{state} \ie\ what the user sees at any given moment,
the set of \emph{commands} available to user and what
they do, and finally how we use a command to \emph{update}
the visualization state. In what follows, for clarity of
exposition, we assume we have a (global) trace:
$\stepg{e_0}{\emptysu}{\bullet}{e_n}{\vsu}{\tr}$, where
$e_0$ and $e_n$ are the \emph{initial} and \emph{final}
expressions respectively.

@@

\input{traversals_new.tex}

>>>>

\subsection{Tracing Semantics}
\label{sec:inter-semant}
%
% \begin{figure*}[t]
% %% \relDescription{Computing Sub-Terms}
% %% \begin{gather*}
% %% \begin{array}{lcl}
% %% \subtermssym                 & \dcolon & e \to \tr \\
% %% \subterms{\eapp{e_1}{e_2}}   & \defeq & \subterm{\eapp{e_1}{e_2}}{e_1}; \subterm{\eapp{e_1}{e_2}}{e_2} \\
% %% \subterms{\eplus{e_1}{e_2}}   & \defeq & \subterm{\eplus{e_1}{e_2}}{e_1}; \subterm{\eplus{e_1}{e_2}}{e_2} \\
% %% \subterms{\eif{e_1}{e_2}{e_3}}   & \defeq & \subterm{\eif{e_1}{e_2}{e_3}}{e_1}; \\
%                                 %% &        & \subterm{\eif{e_1}{e_2}{e_3}}{e_2}; \\
%                                 %% &        & \subterm{\eif{e_1}{e_2}{e_3}}{e_3} \\
% %% \subterms{\elet{x}{e_1}{e_2}}   & \defeq & \subterm{\elet{x}{e_1}{e_2}}{e_1}; \\
%                                 %% &        & \subterm{\elet{x}{e_1}{e_2}}{e_2} \\
% %% \subterms{\efun{x}{e}}       & \defeq & \subterm{\efun{x}{e}}{e} \\
% %% \subterms{e}                 & \defeq & \bullet
% %% \end{array}
% %% \end{gather*}
% \judgementHead{Traced Evaluation}{\stepg{e}{\su}{\tr}{e}{\su}{\tr}}
% \begin{gather*}
% \inference[\recontext]
%   {\stepg{e}{\su}{\tr}{e'}{\su'}{\tr'}}
%   {\stepg{C[e]}{\su}{\tr}{C[e']}{\su'}{\singlestep{C[e]}{C[e']}; \tr'}}
% \\ \\
% \inference[\reappgood]
%   {\pair{\efun{x}{e}}{\su'} = \force{v_1}{\tfun{\thole{}}{\thole{}}}{\su}}
%   {\stepg{\eapp{v_1}{v_2}}{\su}{\tr}
%          {e\sub{x}{v_2}}{\su'}{\singlestep{\eapp{v_1}{v_2}}{e\sub{x}{v_2}}; \tr}}
% \\ \\
% \inference[\reappbad]
%   {\pair{\stuck}{\su'} = \force{v_1}{\tfun{\thole{}}{\thole{}}}{\su}}
%   {\stepg{\eapp{v_1}{v_2}}{\su}{\tr}{\stuck}{\su'}{\singlestep{\eapp{v_1}{v_2}}{\stuck}; \tr}}
% \end{gather*}
% \caption{A selection of the operational semantics from
%   Figure~\ref{fig:operational}, extended to collect a full reduction
%   graph.}
% \label{fig:interactive}
% \end{figure*}

@@

\paragraph{Tracing Semantics}
%
We extend the transition relation (\S~\ref{sec:semantics}) to
collect the set of edges corresponding to the reduction graph.
%
Concretely, we extend the operational semantics to
a relation of the form $\stepg{e}{\vsu}{\tsu}{\tr}{e'}{\vsu'}{\tsu'}{\tr'}$
where $\tr'$ collects the edges of the transition.

\paragraph{Collecting Edges}
%
Next, we describe the general recipe for extending a transition
rule to collect edges.
%
The general recipe for collecting steps-to edges is to
to recording the consequent of each original rule in the
trace. That is, each original judgment \step{e}{\vsu}{\tsu}{e'}{\vsu'}{\tsu'}
becomes \stepg{e}{\vsu}{\tsu}{\tr}{e'}{\vsu'}{\tsu'}{\singlestep{e}{e'}; \tr}.
%
% As the translation is mechanical, we provide a selection of examples
% in Figure~\ref{fig:interactive}.
As the translation is mechanical, we will not discuss it further.

@@

Next, we show how to build a visual interactive debugger
from the traced semantics, by describing the visualization
\emph{state} \ie\ what the user sees at any given moment,
the set of \emph{commands} available to user and what
they do, and finally how we use a command to \emph{update}
the visualization state. In what follows, for clarity of
exposition, we assume we have a (global) trace:
$\stepg{e_0}{\emptysu}{\emptysu}{\bullet}{e_n}{\vsu}{\tsu}{\tr}$, where
$e_0$ and $e_n$ are the \emph{initial} and \emph{final}
expressions respectively.

@@

% \input{traversals_new.tex}
